---
title: "Cov(inc period, serial interval) in Tianjin"
author: "Jessica Stockdale"
date: "May 19, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(viridis)
library(tidyverse)
library(lubridate)
options(digits=3)
set.seed(3456)
```

## Tianjin

First, load in the data

```{r}
tdata=read.csv("data/Tianjin135cases_revised.csv",na.strings = "", stringsAsFactors = F)
tdata$symptom_onset=as.Date(tdata$symptom_onset, format = "%d/%m/%Y")
tdata$start_source=as.Date(tdata$start_source, format = "%d/%m/%Y")
tdata$end_source=as.Date(tdata$end_source,format = "%d/%m/%Y" )
tdata$confirm_date=as.Date(tdata$confirm_date,format = "%d/%m/%Y" )
tdata <- tdata[which(!is.na(tdata$symptom_onset)),]
names(tdata)[1] = "case_id"

tdata$Infection_source <- str_replace(tdata$Infection_source, pattern = "JN", replacement = "TJ")

```

We need to do some preprocessing. NOTE: This step involves assuming that, in each pair, the one who showed symptoms first was the infector. 

```{r}
tdata$Infection_source <- str_to_lower(tdata$Infection_source)
tdata$Infection_source <- str_trim(tdata$Infection_source)
tdata$Infection_source_dup <- tdata$Infection_source
tdata$Infection_source_dup <- str_replace_all(tdata$Infection_source_dup, pattern = "person", replacement = "individual")
tdata$Infection_source_dup <- str_replace(tdata$Infection_source_dup, 
                                          pattern = "coworker of a individual from wuhan",
                                          replacement = "coworker")


tdata <- mutate(tdata, source_group = case_when(!is.na(str_match(Infection_source_dup, "wuhan|hubei")) ~ "Wuhan and Hubei", #Priority 1
                                                  !is.na(str_match(Infection_source_dup, "mall|store|shopper|shopping")) ~ "Mall", #Priority 1
                                                  !is.na(str_match(Infection_source_dup, "family|relative|wife|mother|son|sister|daughter|brother|husband|duaghtor|whife|hunsband")) ~ "Relative", #Priority 2
                                                  !is.na(str_match(Infection_source_dup, "coworker|business|workplace|colleague|colleage")) ~ "Coworker", #Priority 2
                                                  !is.na(str_match(Infection_source_dup, "tj|patient")) ~ "Other relationship", #Priority 2
                                                  !is.na(str_match(Infection_source_dup, "train|travel|trip|hebei|dalian")) ~ "Other travel", #Priority 3
                                                  !is.na(str_match(Infection_source_dup, "unknown|unclear")) ~ "Unknown", #Priority 5
                                                  is.na(Infection_source_dup) ~ "Unknown", #Priority 5
                                                  T ~ "other")) #there should be none of these, so this is just a sanity check!  

#Remove the duplicated infection source column which is no longer necessary
tdata <- select(tdata, -Infection_source_dup)

#What is distribution of probably source of infection (grouped)?
table(tdata$source_group) 

mynodes <- tdata$case_id

mynodes <- str_to_lower(mynodes) 
tdata$case_id <- str_to_lower(tdata$case_id)

edges = data.frame(from=mynodes[9],to=mynodes[21],stringsAsFactors = F ) # i read this one manually 

for (id in 1:nrow(tdata)) {
tonode=tdata$case_id[id]
fromnodes=str_extract_all(tdata$Infection_source[id], "tj\\d+", simplify = T) #in lower case due to above early/late split on infection source
  if (length(fromnodes)>0) {
    for (k in 1:length(fromnodes)) {
      edges=rbind(edges, c(fromnodes[k], tonode))
    }
  }
}
head(edges)
edges=edges[-1,] #Remove the initial relationship we gave so it isn't duplicated
edges=edges[-which(is.na(edges[,1])),] # NAs arose from a few empty entries for Infection_source 

tdata_sympt <- select(tdata, case_id, symptom_onset)

names(tdata_sympt) <- str_replace(names(tdata_sympt), "case_id", "from")
undir_tdates <- left_join(edges, tdata_sympt, by = "from")
names(undir_tdates) <- str_replace(names(undir_tdates), "symptom_onset", "from_sympt_date")

# Repeat, but add the date of symptom onset for the caseID of the 'to' case
names(tdata_sympt) <- str_replace(names(tdata_sympt), "from", "to")
undir_tdates <- left_join(undir_tdates, tdata_sympt, by = "to")
names(undir_tdates) <- str_replace(names(undir_tdates), "symptom_onset", "to_sympt_date")

undir_tdates <- mutate(undir_tdates, earliest_sympt_onset = pmin(to_sympt_date, from_sympt_date, na.rm = T), 
                                   raw_serial_interval = to_sympt_date - from_sympt_date,   
                                   abs_serial_interval = abs(raw_serial_interval))

pos <- filter(undir_tdates, raw_serial_interval >= 0)
neg <- filter(undir_tdates, raw_serial_interval < 0)
onlyone <- filter(undir_tdates, is.na(raw_serial_interval)) 

names(neg)
names(neg)[1] <- "to"
names(neg)[2] <- "from"
names(neg)[3] <- "to_sympt_date"
names(neg)[4] <- "from_sympt_date"
names(neg)

undir_tdates <- bind_rows(pos, neg, onlyone)

undir_tdates$pto <- str_replace(undir_tdates$to, pattern = "tj", replacement = "")
undir_tdates$pto <- str_pad(undir_tdates$pto, width = 3, side = "left", pad = "0")

undir_tdates$pfrom <- str_replace(undir_tdates$from, pattern = "tj", replacement = "")
undir_tdates$pfrom <- str_pad(undir_tdates$pfrom, width = 3, side = "left", pad = "0")

undir_tdates <- mutate(undir_tdates, pairID = factor(paste("tj", pfrom, "-", "tj", pto, sep = "")))

rm(pos, neg, onlyone)

```

We also need to get the incubation periods

```{r}
tdata$end_source[which(is.na(tdata$end_source))]=tdata$symptom_onset[which(is.na(tdata$end_source))]  # if no end exposure: set to symptom onset 
tdata$end_source = pmin(tdata$end_source, tdata$symptom_onset) # if end exposure after onset, set to onset 
tdata$start_source[which(is.na(tdata$start_source))]= tdata$symptom_onset[which(is.na(tdata$start_source))] - 20 # if no start, set to symptom onset - 20

tdata$maxIncTimes=tdata$symptom_onset-tdata$start_source 
tdata$minIncTimes = tdata$symptom_onset-tdata$end_source

tdata$maxIncTimes = pmax(3, tdata$maxIncTimes)
tdata$minIncTimes = pmax(1, tdata$minIncTimes)

```

We want to make a data frame with a row for every suspected infector-infectee pair - and including the serial interval for this pair, and the incubation period of both infector and infectee. 

```{r}

tianjin.data <- data.frame(infector = undir_tdates$from, infectee = undir_tdates$to, serial.interval = undir_tdates$abs_serial_interval, inc.infector.min = tdata$minIncTimes[match(undir_tdates$from,tdata$case_id)], inc.infector.max = tdata$maxIncTimes[match(undir_tdates$from,tdata$case_id)], inc.infectee.min =  tdata$minIncTimes[match(undir_tdates$to,tdata$case_id)], inc.infectee.max = tdata$maxIncTimes[match(undir_tdates$to,tdata$case_id)])

```

There are some NAs at the end of this, which we filter out

```{r}

tianjin.data = tianjin.data[!is.na(tianjin.data$serial.interval),]

```
