---
title: "Cluster Network Data Preprocessing & Plotting"
author: "Emma Garlock; Yen-Hsiang (Brian) Lee"
date: "18/02/2020"
updated: "23/02/2020"
output: html_document
---


You can go through this to see how we got the data ready to plot, or skip to line 91 where you can just upload the final dataset and plot it. 

Load all the libraries we will need 
```{r}
# getwd()
library(ggplot2)
library(tidyverse)
library(ggnetwork)
library(OneR)
library(igraph)
library(here)
library(viridis)
library(paletteer)
```

load up all the links we have (just the first two column of the dataframe expanded)
```{r}
ncov_sing = read.csv("data/COVID-19_Singapore_formated_dates.csv")
```

## You can use the sript below to create a weighted network plot. This allows us to have everything arranged spatially but only maintain the links in the `Related Cases` column when rendering the visualization. 
Trying to make weighted edges 


```{r}
#make a list so that we can supply weights to the edges 
w_edge=ncov[,c(1:3)]#only need these from the pain dataset 
names(w_edge)=c("CaseID","2","1")#change the names 
w_edge[w_edge==""] <- NA# make the blanks into NAs for better sorting 

w_edge_long=pivot_longer(w_edge,cols=c("2","1"),names_to="weight")%>% #pivot longer 
  separate( 'value', paste("relations", 1:22, sep=""), sep=",", extra="drop")%>%#separate out all the connections 
  pivot_longer(cols=relations1:relations22,names_to="relation")%>%#pivot longer again 
  select(CaseID,value,weight)%>%#just get the columns we need 
  filter(value !=" " | !is.na(value) |value !="")%>%#filter out any with missing values 
  filter(weight=="1")#just use weight-1 (these are just to help get things laid out poroperly during plotting, these won't actually show up in the figure. The ones that show up in the figure come later.)

names(w_edge_long)=c("from","to","weight")#change names so things will merge 

#using fedges fom the singapore_serial_intervals_revised.Rmd 
fedges_weighted=fedges%>%select(-arrows)
fedges_weighted$weight=2
fedges_weighted$from=as.character(str_extract_all(fedges_weighted$from,"\\(?[0-9,.]+\\)?"))
fedges_weighted$to=as.character(str_extract_all(fedges_weighted$to,"\\(?[0-9,.]+\\)?"))
plot_edges=rbind(w_edge_long,fedges_weighted)
plot_edges$test=plot_edges$to %in% ncov_sing$CaseID
plot_edges=plot_edges%>%
  filter(test=="TRUE")%>%#remove nodes that don't exist in the ncov_sing due to missing data (they still exist inthe relationships column which is why they showed up)
  select(from,to,weight)#select just the columns we need. 

#write.csv(eg_edges2,"data/sing_edges_eg.csv")
```


```{r}
cov_weighted=graph_from_data_frame(plot_edges, vertices = nodes_sing, directed = FALSE)
cov_net_weight=ggnetwork(cov_weighted)
#write.csv(cov_net_weight,("data/singapore_weighted_network_0504.csv"))
```


```{r}
#cov_net_weight=read.csv(("data/singapore_weighted_network_0504.csv"))
cov_net_w = ggplot(cov_net_weight, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(aes(), subset(cov_net_weight,weight==2),linetype="solid") +
  geom_nodes(aes(color=cluster),size=8) +
  geom_text(aes(label=name),check_overlap = TRUE) +
  geom_text(aes(label=name),subset(cov_net_weight, cluster == "Grace Assembly of God"),colour="white",check_overlap = TRUE) +
  guides(colour = guide_legend(title.position="top", title.hjust = 0.5, override.aes = list(size=3))) +
  scale_colour_viridis_d(na.value="lightgrey") +
  labs(color="Cluster")+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) + #centre main title
  theme(panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.text=element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal")
cov_net_w

ggsave("final_figures/Fig4_singapore_cov_network_revised_0505.pdf",plot=cov_net_w,device="pdf",height=8,width=11,units="in")
```


####Tianjin Data####

There won't be any weighting or anything with the Tianjin one so we can just use the same edge list that was made in tianjin_serial_intervals_revised.Rmd 
- tedges 
- nodes 
 But we still have to filter based on nodes that were removed based on missing data 
```{r}
tedges$test=tedges$from %in% nodes$id
tedges=tedges%>%filter(test=="TRUE")%>%select(-test)
```
 

Make the dataset for plotting , and make a new column that is just the number without the "TJ" to make for better node label 
```{r}
tj_net=graph_from_data_frame(tedges,vertices = nodes,directed = FALSE)
ntj=ggnetwork(tj_net)

ntj$case_num=as.character(str_extract_all(ntj$name,"\\(?[0-9,.]+\\)?"))
#write.csv(ntj_clean,here("network_diagram/tianjin_cluster_known_edges.csv"))

```
Make the plot 
```{r}
#Upload the tianjin_cluster_known_edges.csv for this plot 
#ntj_clean=read.csv(here("network_diagram/tianjin_cluster_known_edges.csv"))
tj_cov_net = ggplot(ntj, aes(x = x, y = y, xend = xend, yend = yend))+
  geom_edges() +
  geom_nodes(aes(color=group),size=8) +
  geom_text(aes(label=case_num), check_overlap = TRUE)+
  #geom_text(aes(label=case_num),subset(ntj,cluster==" close contact "),colour="white",check_overlap = TRUE)+
  scale_color_viridis_d()+
  labs(color="Cluster")+
  guides(colour = guide_legend(title.position="top", title.hjust = 0.5, override.aes = list(size=3))) +
  #ggtitle("Clusters of COVID-19 cases in Tianjin, China")+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.text=element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal")


tj_cov_net

ggsave("final_figures/Fig4_tianjin_cov_network_revised_0505.pdf",plot=tj_cov_net,device="pdf",height=8,width=11,units="in")
```




