---
title: "Cluster Network Data Preprocessing & Plotting"
author: "Emma Garlock; Yen-Hsiang (Brian) Lee"
date: "18/02/2020"
updated: "23/02/2020"
output: html_document
---


You can go through this to see how we got the data ready to plot, or skip to line 91 where you can just upload the final dataset and plot it. 

Load all the libraries we will need 
```{r}
# getwd()
library(ggplot2)
library(tidyverse)
library(ggnetwork)
library(OneR)
library(igraph)
library(here)
library(viridis)
library(paletteer)
```

load up all the links we have (just the first two column of the dataframe expanded)
```{r}
ncov_sing = read.csv("data/COVID-19_Singapore_formated_dates.csv")
```

## You can use the sript below to create a weighted network plot. This allows us to have everything arranged spatially but only maintain the links in the `Related Cases` column when rendering the visualization. 
Trying to make weighted edges 


```{r}
w_edge=ncov[,c(1:3)]
names(w_edge)=c("CaseID","2","1")
w_edge[w_edge==""] <- NA
w_edge_long=pivot_longer(w_edge,cols=c("2","1"),names_to="weight")%>% 
  separate( 'value', paste("relations", 1:22, sep=""), sep=",", extra="drop")%>%
  pivot_longer(cols=relations1:relations22,names_to="relation")%>%
  select(CaseID,value,weight)%>%
  filter(value !=" " | !is.na(value) |value !="")

names(w_edge_long)=c("from","to","weight")
w_edge_long$to=gsub("^\\s+|\\s+$", "",w_edge_long$to)
w_edge_long=w_edge_long[w_edge_long$from !="53",]

eg_edges=w_edge_long%>%
  filter(weight==2)%>%
  select(-weight)%>%
  distinct()

eg_edges2=eg_edges[!duplicated(data.frame(t(apply(eg_edges,1,sort)))),]

write.csv(eg_edges2,"data/sing_edges_eg.csv")
```


```{r}
cov_weighted=graph_from_data_frame(w_edge_long, vertices = nodes_sing, directed = FALSE)
cov_net_weight=ggnetwork(cov_weighted)
write.csv(cov_net_weight,("data/singapore_weighted_network_0504.csv"))
```


```{r}
cov_net_weight=read.csv(("data/singapore_weighted_network_0504.csv"))
cov_net_w = ggplot(cov_net_weight, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(aes(), subset(cov_net_weight,weight==2),linetype="solid") +
  geom_nodes(aes(color=cluster),size=8) +
  geom_text(aes(label=name),check_overlap = TRUE) +
  geom_text(aes(label=name),subset(cov_net_weight, cluster == "Grace Assembly of God"),colour="white",check_overlap = TRUE) +
  guides(colour = guide_legend(title.position="top", title.hjust = 0.5, override.aes = list(size=3))) +
  scale_colour_viridis_d(na.value="lightgrey") +
  labs(color="Cluster")+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) + #centre main title
  theme(panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.text=element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal")
cov_net_w

ggsave("final_figures/Fig4_singapore_cov_network_revised_0504.pdf",plot=cov_net_w,device="pdf",height=8,width=11,units="in")
```


####Tianjin Data####

Using the edges from the Tianjin.Rmd cluster-load the data from that script. 

You can go through this to see how we got the data ready to plot, or skip to line 164 where you can just upload the final dataset and plot it. 

```{r}

tdata <- read.csv("data/Tianjin135casesFeb22.csv",na.strings = "", stringsAsFactors = F)

tdata$symptom_onset=as.Date(tdata$symptom_onset, format = "%d/%m/%Y")
tdata$start_source=as.Date(tdata$start_source, format = "%d/%m/%Y")
tdata$end_source=as.Date(tdata$end_source,format = "%d/%m/%Y" )
tdata$confirm_date=as.Date(tdata$confirm_date,format = "%d/%m/%Y" )

mynodes <- tdata$case_id

#Transform everything to lower case to make sure there aren't any issues with matching due to case inconsistencies
mynodes <- str_to_lower(mynodes) 
tdata$case_id <- str_to_lower(tdata$case_id)

edges = data.frame(from=mynodes[9],to=mynodes[21],stringsAsFactors = F ) # i read this one manually 

for (id in 1:nrow(tdata)) {
tonode=tdata$case_id[id]
fromnodes=str_extract_all(tdata$Infection_source[id], "tj\\d+", simplify = T) #in lower case due to above early/late split on infection source
  if (length(fromnodes)>0) {
    for (k in 1:length(fromnodes)) {
      edges=rbind(edges, c(fromnodes[k], tonode))
    }
  }
}
head(edges)
edges=edges[-1,] #Remove the initial relationship we gave so it isn't duplicated
edges=edges[-which(is.na(edges[,1])),] # NAs arose from a few empty entries for Infection_source 
edges_dist = select(edges, from, to)
edges_dist = distinct(edges_dist)
```


Make the dataset for plotting , and make a new column that is just the number without the "TJ" to make for better node label 
```{r}
tj_net=graph_from_data_frame(edges_dist,vertices = tdata,directed = FALSE)
ntj=ggnetwork(tj_net)

ntj$case_num=as.character(str_extract_all(ntj$name,"\\(?[0-9,.]+\\)?"))
#write.csv(ntj_clean,here("network_diagram/tianjin_cluster_known_edges.csv"))

```
Make the plot 
```{r}
#Upload the tianjin_cluster_known_edges.csv for this plot 
#ntj_clean=read.csv(here("network_diagram/tianjin_cluster_known_edges.csv"))
tj_cov_net = ggplot(ntj, aes(x = x, y = y, xend = xend, yend = yend))+
  geom_edges() +
  geom_nodes(aes(color=source_group),size=8) +
  geom_text(aes(label=case_num), check_overlap = TRUE)+
  #geom_text(aes(label=case_num),subset(ntj,cluster==" close contact "),colour="white",check_overlap = TRUE)+
  scale_color_viridis_d()+
  labs(color="Cluster")+
  guides(colour = guide_legend(title.position="top", title.hjust = 0.5, override.aes = list(size=3))) +
  #ggtitle("Clusters of COVID-19 cases in Tianjin, China")+
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.text=element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal")


tj_cov_net

#ggsave("../final_figures/Fig4b_tj_static_network.pdf",plot=tj_cov_net,device="pdf",path=here("network_diagram"),height=8,width=11,units="in")
```
Ideally I can get the network diagram to cluster accoring to the broader clusters and not jsut based on the known conncections. It doesn't change the information, just shows it in a much nicer way. So we won't use the list of edges generated above, but we will use that list to confirm the accuracy of the new diagram. 

Extract the known contacts from the large dataset-same procedure used above. Need to clean up some of the cells, removing white spaces and such (196)
```{r}
#tj=read.csv(here("Clustering/Tianjin135casesFeb22.csv"))
edges_unclean2=tdata
edges_unclean2$close_conn=str_extract_all(edges_unclean2$Infection_source, regex("TJ\\d+"))
edges_unclean2$close_conn=ifelse(edges_unclean2$close_conn=="character(0","NA",as.character(edges_unclean2$close_conn))
edges_unclean2$close_conn=gsub("^c\\(|\\)$", "", edges_unclean2$close_conn)

head(edges_unclean2)
#write.csv(edges_unclean,here("Clustering/tianjin_edges.csv"))
```

Need to clean up the cluster names (207-209) and make al ist of the known contacts (210) then merge them with the full datatset so that we can include the cleaned cluster names (after this block called infection_source_named)

```{r}
tdata_cc=read.csv(here("tianjin_edges_clean.csv"))
colnames(tdata_cc)[4]="cluster"
edges_cleaned=tdata_cc
edges_cleaned$cluster=str_replace(edges_cleaned$cluster,"coworker ","coworker")
edges_cleaned$cluster=str_replace(edges_cleaned$cluster,"confirmed patient","unknown")
edges_cleaned$cluster=str_replace(edges_cleaned$cluster,"not specified","unknown")
edges_cleaned$edge_list=paste(paste(edges_cleaned$edge1,edges_cleaned$edge2,edges_cleaned$edge3,edges_cleaned$edge4,sep=","))


edges_full_df=merge(edges_cleaned,edges_unclean2,by="case_id")
edges_full=edges_full_df[,c(1,4,22)]
colnames(edges_full)[2]="infection_source_named"


```

make a list of all the fifferent clusters and make a bunch of lists of all the different cases located within those clusters 
```{r}
cluster_list=(unique(edges_cleaned$cluster))
for (c in cluster_list){
  assign(paste0("list",c),edges_cleaned[edges_cleaned$cluster==c,2])
}


```
assign a cluster list (made above) to each row based on what cluster they belong to 
```{r}

edges_full$cluster_list=ifelse(edges_full$infection_source_named=="Wuhan",list(as.character(listWuhan)),
                            ifelse(edges_full$infection_source_named=="Wuhan; train import",list(as.character(`listWuhan; train import`)),
                            ifelse(edges_full$infection_source_named=="unkown","NA",
                            ifelse(edges_full$infection_source_named=="travel",list(as.character(`listtravel `)),
                            ifelse(edges_full$infection_source_named=="train",list(as.character(listtrain)),
                            ifelse(edges_full$infection_source_named=="mall",list(as.character(listmall)),
                            ifelse(edges_full$infection_source_named=="Hubei",list(as.character(listHubei)), 
                            ifelse(edges_full$infection_source_named=="coworker",list(as.character(listcoworker)),
                            ifelse(edges_full$infection_source_named==" close contact ","NA","NA")))))))))
                                
```
Clean up the lists since they have brackets and quotation marks and other strange characters or things that will mess up the analysis 
```{r}

edges_reduced=edges_full
names(edges_reduced)=c("case_id","infection_source_named","edge_list","cluster_list")
edges_reduced$cluster_list=ifelse(edges_reduced$cluster_list=="NA",as.character(edges_reduced$edge_list),as.character(edges_reduced$cluster_list))
edges_reduced$cluster_list=as.character(edges_reduced$cluster_list)
edges_reduced$cluster_list=gsub("^c\\(|\\)$", "", edges_reduced$cluster_list)
edges_reduced$cluster_list=gsub('"', "", edges_reduced$cluster_list)
```
separate each list so that we can convert it to long format, then replace the character(0 that has been hanging around 
```{r}
edges_long=separate(edges_reduced, 'cluster_list', paste("relations", 1:31, sep=""),sep=",",extra="drop")%>%select(-edge_list)
edges_exp_long_na=gather(edges_long, rel, link, relations1:relations31)%>%select(-rel)
edges_exp_long_na$link=ifelse(edges_exp_long_na$link=="character(0","NA",as.character(edges_exp_long_na$link))
```
Remove the NAs, blaks and character "NA,  subset to get the edge list, remove white spaces, and remove and duplicate edges 
```{r}
edges_exp_long = edges_exp_long_na[!(is.na(edges_exp_long_na$link) | edges_exp_long_na$link == "NA" | edges_exp_long_na$link == ""),]
edges_exp_long=edges_exp_long[,c(1,3)]
names(edges_exp_long)=c("from","to")
edges_exp_long=as.data.frame(apply(edges_exp_long,2,function(x)gsub('\\s+', '',x)))
edges_exp_long[1:2] <- t(apply(edges_exp_long[1:2], 1, sort))
edges_dist_full = distinct(edges_exp_long)

```
            
make the dataset for network plotting, but remove the first column because that is just an index column. also make a new coumn that is just the case number without the "TJ" so that it plots better as a node label        
```{r}
tdata_cc=tdata_cc[,-1]
tj_net_ic=graph_from_data_frame(edges_dist_full,vertices = edges_full_df,directed = FALSE)
ntj_ic=ggnetwork(tj_net_ic)

ntj_ic$case_num=str_extract_all(ntj_ic$name,"\\(?[0-9,.]+\\)?")
ntj_ic$edge1 <- gsub('\\s+', '', ntj_ic$edge1)


#write.csv(ntj_ic,here("Clustering/network_diagram/tianjin_cluster_all_edges.csv"))

```
Make the plot, try and have only the edges that are known connections showing. 
```{r}
#upload tianjin_cluster_all_edges.csv for this plot
tj_cov_net_ic = ggplot(ntj_ic, aes(x = x, y = y, xend = xend, yend = yend))+
  geom_edges(aes(),subset(ntj_ic,edge_list!=",,,"),linetype="solid",colour="black") +
  geom_nodes(aes(color=cluster),size=8) +
  geom_text(aes(label=case_num))+
  geom_text(aes(label=case_num), subset(ntj_ic,cluster==" close contact "),colour="white")+
  scale_color_viridis_d()+
  ggtitle("Clusters of COVID-19 cases in Tianjin, China")+
  theme_blank()

tj_cov_net_ic

#ggsave("tianjin_cluster_feb28v1.png",plot=tj_cov_net_ic,device="png",path=here("Clustering/network_diagram"))
```

This is not wokring yet, when double checking wiht the egdes_dist list from above, we are still not having some show up and the network is not being made properly 

