---
title: "tianjian_models"
author: "Michelle Coombe"
date: "18/02/2020"
output: html_document:
  keep_md: TRUE
---

## Modeling initial Tianjin data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(deSolve)
library(ggplot2)
library(tidyverse)
```

## Introduction

This is an implementation and exploration of the susceptible-exposed-infectious (SEI) model in Wu et al, Lancet, https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(20)30260-9/fulltext to the Tianjin outbreak of the 2019 novel corona virus.


## Data

From Data folder in Hackathon2020 git; 'Tianjin 125 cases' from Feb 17th


```{r}
tj_data <-read_csv("tianjin_data/Tianjin_125_cases_updated_Feb_17_csv.csv")
head(tj_data)
```

Change dates to date format and then turn into a numeric day value since first day of local outbreak in Tianjin 
i.e. first day any confirmed or symptoms in a later confirmed case were detected

*TODO* look into which start date to use; similar to what CC did or similar to Wu paper; does including a longer tail of zeros change the model results?

```{r}
tj_data$symptom_onset <- as.Date(tj_data$symptom_onset, format = "%d/%m/%Y")
tj_data$confirm_date <- as.Date(tj_data$confirm_date, format = "%d/%m/%Y")

min(min(tj_data$symptom_onset, na.rm = T), min(tj_data$confirm_date, na.rm = T))

# Adding a new column to provide the numeric day since the outbreak started in Tianjin
tj_data <- mutate(tj_data, sympt_days_since_outbreak = as.numeric(symptom_onset - min(min(tj_data$symptom_onset, na.rm = T), min(tj_data$confirm_date, na.rm = T))),
                  confirm_days_since_outbreak = as.numeric(confirm_date - min(min(tj_data$symptom_onset, na.rm = T), min(tj_data$confirm_date, na.rm = T))))
```

Group number of cases per day since start of local outbreak 
*NOTE* this automatically removes missing data if no date of symptom onset
*TODO* consider imputing data using confirmation date MINUS mean incubation period; should be pretty straight forward using mutate and an if_else statement (i.e. if value is not na, use that, otherwise impute as per formula)

```{r}
tj_final = data.frame(day=0:max(tj_data$sympt_days_since_outbreak, na.rm = T),  xd = vapply(0:max(tj_data$sympt_days_since_outbreak, na.rm = T), 
  function(x) sum(tj_data$sympt_days_since_outbreak == x, na.rm = TRUE),
 FUN.VALUE = 1)) 
```


## Summary statistics to use for stratification of the model
Change gender to a factor with labels, rather than 0 vs 1, which are male vs female (respectively), as per datasheet
Not much discrepancies between age or sex groups, so maybe let's look at different possible sources for the 

```{r}
tj_data$gender <- factor(tj_data$gender,
                         levels = c(0, 1),
                         labels = c("male", "female"))


```

## ODE Model specification

We need time-varying input parameters to account for the introductions $z(t)$ and the chunyun festival adjustments for travel information. Note that I have NOT implemented control measures -- something to be added! 

*TODO:* look up if we need to account for the chunyun festival in singapore? probably because there could be more movement between mainland China and Singapore during this time period

The ODE model takes in the usual time, state and parameters, but also two functions. These allow parameters to vary with time. 

zinput is the force of zoonotic infections, so will be zero for Tianjin and Singapore models

```{r}
tj_model <- function(t, state, pars, zinput, chuinput) {
  with(c(as.list(state), pars), {     # make it so we can use variable and par names
    zt = zinput(t)
    LcwT= Lcw + chuinput(t)*(LcwCH-Lcw)
    LwcT= Lwc + chuinput(t)*(LwcCH-Lwc)
#    LcwT = LcwT*travelban(t) # 1 or 0 # we should add the travelban, control measures 
#    LwcT = LwcT*travelban(t) # 1 or 0 # as anotther time-dep function
    dSdt = -(S/N)*((R0/DI)*I+ zt) + Liw + LcwT - (1/N)*(Lwi+LwcT)*S
    dEdt = (S/N)*((R0/DI)*I+ zt) -E/DL - (1/N)*(Lwi+LwcT)*E
    dIdt = E/DL - I/DI - (1/N)*(Lwi+LwcT)*I
    list(c(dSdt, dEdt, dIdt))
  })
}
```

Now specify the starting state, the time range, the parameters and the input functions. It seems that Wu et al assumed a total of 86 introductions from Dec 1 to Dec 31 when the market was closed, so we model a rate of 86/31 introductions per day. 

For Tianjin:
zt = 0 because no source of zoonotic infections (only human to human)
S = Population Tianjin *TODO* look up!!!
E = 0
I = don't hard code, but will depend on data because the first cases are human
  *TOOO* confirm this!!! also; should we have a "force of human infection" with a factor indicating that more cases are coming in from other international travelers from Wuhan (ie need to times by # of infected people in Wuhan/ number of people in Wuhan and subtract from the dI/dt equation)

