---
title: "Incidence plots"
author: "Michelle Coombe"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages and data
```{r}
library(tidyverse)
library(lubridate)
library(incidence)
#library(RColorBrewer)
#library(viridis)
library(mgcv)

s.data <- read_csv("data/COVID-19_Singapore - singapore_ncov_2019_fix.csv")
t.data <- read_csv("Tianjin135casesFeb22.csv")

#Need to replace one date which is labelled as "none" with NA
t.data$symptom_onset[which(t.data$symptom_onset == "none")] <- NA
```

## Select variables of interest and clean for incidence plotting purposes
There are two possible choices of dates we can use to demonstrate the incidence curve of the epidemic: the date of symptom onset or the date of COVID-19 confirmation. The date of symptom onset is biologically more consistent (there is nothing about the nature of the virus that would mean each case would be confirmed on the same day post-infection); however, there is missing data for 10 cases (in both Tianjin and Singapore datasets). Therefore, I will make the curves using the date of symptom onset, but we could also run the same analysis using the date of confirmation instead as a sensitivity analysis to make sure it does not impact the results.


```{r}
### 1. Singapore dataset
head(s.data$date_onset_symptoms)
sum(is.na(s.data$date_onset_symptoms)) #10

# Make a column for grouping, based on the reason why the case was exposed
# Turn 'presumed_reason' into lower case and get trim any whitespace so don't have issues with case sensitivity, etc
s.data$presumed_reason <- str_to_lower(s.data$presumed_reason)
s.data$presumed_reason <- str_trim(s.data$presumed_reason)
table(s.data$presumed_reason)
sum(is.na(s.data$presumed_reason)) #15 NAs

# Make a new column where we group the 'presumed_reason' under a label (known relationship, gathering, wuhan travel) for each of the above three groups
s.data <- mutate(s.data, presumed_reason_group = case_when(!is.na(str_match(presumed_reason, "symptom onset|via")) ~ "known relationship",
                                                           !is.na(str_match(presumed_reason, "wuhan|airport")) ~ "wuhan travel", #'airport' case (CaseID 17) does not have 'wuhan' in reason but does have it under 'Case' column that they are from Wuhan
                                                           !is.na(str_match(presumed_reason, "grace")) ~ "grace assembly of god",
                                                           !is.na(str_match(presumed_reason, "grand")) ~ "grand hyatt singapore",
                                                           !is.na(str_match(presumed_reason, "life")) ~ "life church",
                                                           !is.na(str_match(presumed_reason, "seletar")) ~ "seletar aerospace heights",
                                                           !is.na(str_match(presumed_reason, "yong")) ~ "yong thai hang",
                                                           is.na(presumed_reason) ~ "unknown",
                                                           TRUE ~ "other")) #should not be any other, so is just a double check this has run correctly, especially as dataset grows

# Make a second new column that puts all the gatherings together, possibly for a clearer plot?
s.data <- mutate(s.data, presumed_reason_group2 = case_when(!is.na(str_match(presumed_reason, "symptom onset|via")) ~ "known relationship",
                                                            !is.na(str_match(presumed_reason, "wuhan|airport")) ~ "wuhan travel", #'airport' case (CaseID 17) does not have 'wuhan' in reason but does have it under 'Case' column that they are from Wuhan
                                                           !is.na(str_match(presumed_reason, "grace|grand|life|seletar|yong")) ~ "gatherings",
                                                           is.na(presumed_reason) ~ "unknown",
                                                           TRUE ~ "other"))
table(s.data$presumed_reason_group)
table(s.data$presumed_reason_group2)

# Now select down to columns of interest and turn dates into Date class
s.sympt <- s.data %>% 
           select(c(CaseID, age, sex, 
                    date_onset_symptoms,
                    #date_confirmation,
                    presumed_reason_group,
                    presumed_reason_group2)) %>% 
           mutate(sex = factor(sex),
                  date_onset_symptoms = dmy(date_onset_symptoms),
                  #date_confirmation = dmy(date_confirmation),
                  presumed_reason_group = factor(presumed_reason_group, 
                                                 levels = c("known relationship",
                                                            "wuhan travel",
                                                            "grace assembly of god",
                                                            "grand hyatt singapore",
                                                            "life church",
                                                            "seletar aerospace heights",
                                                            "yong thai hang",
                                                            "unknown")),
                  presumed_reason_group2 = factor(presumed_reason_group2, 
                                                 levels = c("known relationship",
                                                            "wuhan travel",
                                                            "gatherings",
                                                            "unknown")))

# Get rid of rows where date of symptom onset is not known; could be imputed as another option
s.sympt <- s.sympt[which(!is.na(s.sympt$date_onset_symptoms)), ]


### 2. Tianjin dataset
head(t.data$symptom_onset)
sum(is.na(t.data$symptom_onset)) #10

# Make a column for sensible groupings for Tianjin, based on the reason why the case was exposed
# Turn 'Infection_source' into lower case and get trim any whitespace so don't have issues with case sensitivity, etc
t.data$Infection_source <- str_to_lower(t.data$Infection_source)
t.data$Infection_source <- str_trim(t.data$Infection_source)
table(t.data$Infection_source)
sum(is.na(t.data$Infection_source)) #3 NAs

#Note that the order the data are selected in is VERY important to which case goes into which source_group category
  #For those that meet multiple criteria (e.g. wuhan; tj1), the str_match which is highest in the case_when call (i.e. "tj|coworker|patient|partner") will have priority over those matching later 
  #so that the 'source' column contain "wuhan; tj1" would be labelled as infection from a "known relationship" rather than from a "wuhan" origin 

t.data <- mutate(t.data, source_group = case_when(!is.na(str_match(Infection_source, "tj|coworker|patient|partner")) ~ "known relationship", #placing this first means that an individual that has another case listed as source,regardless if there is another reason (i.e. wuhan; tj1) will have 'known relationship' as their source_group reason
                                                  !is.na(str_match(Infection_source, "wuhan")) ~ "wuhan origin",
                                                  !is.na(str_match(Infection_source, "hubei|hebei|dalian")) ~ "other China travel",
                                                  !is.na(str_match(Infection_source, "train|travel")) ~ "location unclear travel",
                                                  !is.na(str_match(Infection_source, "mall|store|shopper")) ~ "mall",
                                                  !is.na(str_match(Infection_source, "unknown|unclear")) ~ "unknown",
                                                  is.na(Infection_source) ~ "unknown",
                                                  T ~ "other")) #there should be none of these, so this is just a sanity check!  
table(t.data$source_group)

# Now select down to columns of interest and turn dates into Date class
t.sympt <- t.data %>% 
          select(case_id, gender, age,
                 symptom_onset,
                 #confirm_date,
                 Infection_source,
                 source_group,
                 severity) %>% 
          mutate(gender = factor(gender),
                 symptom_onset = dmy(symptom_onset),
                 #confirm_date = dmy(confirm_date),
                 source_group = factor(source_group,
                                       levels = c("known relationship",
                                                "wuhan origin",
                                                "other China travel",
                                                "location unclear travel",
                                                "mall",
                                                "unknown")),
                 severity = factor(severity,
                                   levels = c("minor",
                                              "normal",
                                              "severe",
                                              "unclassified")))

# Get rid of rows where date of symptom onset is not known; could be imputed as another option
t.sympt <- t.sympt[which(!is.na(t.sympt$symptom_onset)), ]
```

## Plot daily incidence curves for both clusters
Now to make the incidence plots using the R package 'incidence'. First, we will plot all the weekly cases in each cluster, without any type of grouping. Our cluster outbreaks have only been going for about 1 month so plotting daily incidence makes more sense.
```{r}
s.daily <- incidence(s.sympt$date_onset_symptoms, 
                     interval = 1) 
s.daily
plot(s.daily)

t.daily <- incidence(t.sympt$symptom_onset,
                     interval = 1)
t.daily
plot(t.daily)
```

## Plot daily incidence curves grouped by source of infection, for both clusters
What happens when we look at the incidence plots for each source of infection?
```{r}
### 1. Singapore
s.grouped <- incidence(s.sympt$date_onset_symptoms,
                       interval = 1,
                       groups = s.sympt$presumed_reason_group)
s.grouped
plot(s.grouped,
    # stacked = TRUE,
     border = "grey")
  #Hmmm, kinda hard to read with so many groups

s.grouped2 <- incidence(s.sympt$date_onset_symptoms,
                       interval = 1,
                       groups = s.sympt$presumed_reason_group2)
s.grouped2
plot(s.grouped2,
#     stacked = TRUE,
     border = "grey")
  #Better!

#Hmmm...also kinda hard to read, need to fill each date with NAs for each group in 
#ggplot(s.sympt, aes(x = date_onset_symptoms, fill = presumed_reason_group2)) +
#  geom_bar(position = "dodge") +
#  scale_fill_viridis_d()

### 2. Tianjin
t.grouped <- incidence(t.sympt$symptom_onset,
                       interval = 1,
                       groups = t.sympt$source_group)
t.grouped
plot(t.grouped,
    # stacked = TRUE,
     border = "grey")

```


## Predictions of incidence
We would like to know what the incidence might be in the future, which can be done by using two log-linear regressions of the incidence data (excluding zeros), where the first regression is fit to the growth phase of the epidemic and a second regression model is fit the decreasing phase of the epidemic. This can be done using the 'incidence' package, which will also automatically determine what is the date of the peak of the epidemic (and thus where to split the two regressions)

Another option would be to use a generalized additive model (GAM) to fit the curve of the data: would need to group by date, count # of cases, and I think put in zeros for the dates that had no cases. Could additionally do this by infection source groups.

```{r}
### 1. Singapore
s.fit <- fit_optim_split(s.daily)
s.fit 
s.fit$split # I belive this is the date where the two logistic regressions are split between 'growth' and 'decrease' 
range(s.sympt$date_onset_symptoms)
plot(s.daily, fit = s.fit$fit)

### 2. Tianjin
t.fit <- fit_optim_split(t.daily)
t.fit
t.fit$split  
range(t.sympt$symptom_onset)
plot(t.daily, fit = t.fit$fit)
```

Interestingly, both models have the same split date for the optimal fit...possibly because the date range of cases is very similar? Also, the incidence does seem to be tailing off in both areas; however, it is possible this is a sampling artifact, as perhaps we have not yet found/diagnosed all the individuals that have come down with symptoms during the last week or so. This might be particularly true of the Singapore data where the daily incidence seems to be a little more steady than the daily incidence in Tianjin, which does seem to be showing a more pronounced decrease in daily incidence rate. 

Alternatively, we may be seeing the impact of the travel ban that is leading to tappering numbers of cases, as the Wuhan origin cases are primarily at the start of the epidemic curves for both locations.
