---
title: "Singapore"
author: "Caroline Colijn and Michelle Coombe"
date: "25/02/2020"
output: 
  html_document:
    keep_md = TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survminer)
library(survival)
library(tidyverse)
library(lubridate)
library(icenReg)
library(igraph)
library(visNetwork)
library(stringr)
options(digits=3)

```

## Data 

Thanks to EpiCoronaHack Cluster team. These data are manually entered from postings from the Government of Singapore website:

* source1: TO CONFIRM WITH LAUREN


```{r}
spdata <- read_csv("data/COVID-19_Singapore - singapore_ncov_2019_fix.csv")

# Ensure properly imported
glimpse(spdata)
head(spdata)
tail(spdata)
colSums(is.na(spdata))

# Rename columns 2, 3 and 4 so no spaces
spdata <- rename(spdata, related_cases = starts_with("Related"),
                         cluster_links = "Cluster links",
                         relationship_notes = starts_with("Relation"))

# Change date columns into date objects
spdata <- mutate(spdata, presumed_infected_date = dmy(presumed_infected_date),
                         last_poss_exposure = dmy(last_poss_exposure),
                         symp_presumed_infector = dmy(symp_presumed_infector),
                         date_onset_symptoms = dmy(date_onset_symptoms),
                         date_quarantine = dmy(date_quarantine),
                         date_hospital = dmy(date_hospital),
                         date_confirmation = dmy(date_confirmation),
                         date_discharge = dmy(date_discharge))

# make sure dates parsed properly
range(spdata$presumed_infected_date, na.rm = T)
range(spdata$last_poss_exposure, na.rm = T)
range(spdata$symp_presumed_infector, na.rm = T)
range(spdata$date_onset_symptoms, na.rm = T)
range(spdata$date_quarantine, na.rm = T)
range(spdata$date_hospital, na.rm = T)
range(spdata$date_confirmation, na.rm = T)
range(spdata$date_discharge, na.rm = T)

# Note that the date of symp_presumed_infector for CaseID 79 changed was originally listed as 2020-02-07 (based on online visualizations) but was changed to 2020-02-10, due to Feb 10, 2020 being on the earliest date of onset of symptoms from case 72, as from online info provided, presumed infective contact for CaseID 79 is from 72 (family member), rather than directly from case 52
  singapore$symp_presumed_infector[singapore$CaseID == 79] <- ymd("2020-02-10")
```



## Incubation period

The incubation period is the time between exposure and the onset of symptoms. We estimate this directly from the stated start and end times for cases' exposure windows. Because it is explicitly about the symptom onset, we remove those who don't have symptom onset defined. These are a small minority of cases and the alternative would be to impute their symptom onset time using the others' delay to confirmation time.  For now, we remove them.  

Then, if no other end time for the exposure is given or if the end of the exposure time is after the time of symptom onset, set the last exposure time to the symptom onset time. This is because they must have been exposed before symptom onset.  

If no other start time is given, they must have been exposed since the start of the outbreak (Dec 1, 2019). These give us the maximum and minimum incubation times.  

```{r}
# Note that the date of symp_presumed_infector for CaseID 79 changed was originally listed as 2020-02-07 (based on online visualizations) but was changed to 2020-02-10, due to Feb 10, 2020 being on the earliest date of onset of symptoms from case 72, as from online info provided, presumed infective contact for CaseID 79 is from 72 (family member), rather than directly from case 52

# Change symp_presumed_infector to Feb 10, 2020 (date of symptom onset from caseID 72, the presumed infector)
spdata$symp_presumed_infector[spdata$CaseID == 79] <- ymd("2020-02-10")

# havesymp <- which(!is.na(spdata$date_onset_symptoms)) #index of cases that do not have info on date of symptom onset ; not sure I need this???

spdata <- filter(spdata, !is.na(date_onset_symptoms)) #Remove all the cases that do not have info on date of symptom onset

spdata$last_poss_exposure[which(is.na(spdata$last_poss_exposure))] <- spdata$date_onset_symptoms[which(is.na(spdata$last_poss_exposure))]  # if no end exposure: set to symptom onset 
spdata$last_poss_exposure <- pmin(spdata$last_poss_exposure, spdata$date_onset_symptoms) # if end exposure after onset, set to onset 

spdata$presumed_infected_date[which(is.na(spdata$presumed_infected_date))] <- ymd("2019-12-31") # start date 

sum(is.na(spdata$last_poss_exposure))  
sum(is.na(spdata$presumed_infected_date))

# Make new columns for max and min incubation dates
spdata$maxIncTimes <- spdata$date_onset_symptoms - spdata$presumed_infected_date 
spdata$minIncTimes <- spdata$date_onset_symptoms - spdata$last_poss_exposure

spdata$maxIncTimes
spdata$minIncTimes
```

We use survival analysis in the icenReg package to make parametric estimates, and we use the regular survival package to estimate the time to onset of symptoms. 

```{r}
ggsurvplot(
  fit <- survfit(Surv(spdata$minIncTimes, spdata$maxIncTimes, type="interval2") ~ 1, data = spdata), 
  xlab="Days",
  ylab = "Overall probability of no symptoms yet")
```

For a parametric estimate we remove any remaining NAs and use interval censoring, because we know only that exposure was some time between the minimum and maximum possible values. 

```{r}
sum(is.na(spdata$minIncTimes)) # 0

#Error in model fit; uncensored times cannot be equal to 0; so try replacing exact times = 0 with really small numbers
nozerodata <- spdata
table(nozerodata$minIncTimes)
nozerodata$minIncTimes[nozerodata$minIncTimes == 0] <- 0.0001
table(nozerodata$minIncTimes)

table(nozerodata$maxIncTimes)
nozerodata$maxIncTimes[nozerodata$maxIncTimes == 0] <- 0.0001
table(nozerodata$maxIncTimes)

spfit <- ic_par(Surv(nozerodata$minIncTimes, nozerodata$maxIncTimes, type="interval2") ~ 1, data = nozerodata, dist = "weibull")
summary(spfit)
```

The log of the shape parameter is `r myfit$coefficients[1]` $\pm$ `r sqrt(myfit$var[1,1])`, which gives a shape parameter of `r exp(myfit$coefficients[1])` with a 1.96-sd (in the log) giving the range (`r exp(myfit$coefficients[1]-1.96*sqrt(myfit$var[1,1]))`, `r exp(myfit$coefficients[1]+1.96*sqrt(myfit$var[1,1]))`).

Similarly the log scale parameter is `r myfit$coefficients[2]` $\pm$ `r sqrt(myfit$var[2,2])`, which gives a scale parameter of `r exp(myfit$coefficients[2])` with a one-sd (in the log) giving the range (`r exp(myfit$coefficients[2]-1.96*sqrt(myfit$var[2,2]))`, `r exp(myfit$coefficients[2]+1.96*sqrt(myfit$var[2,2]))`). 

```{r}
interqs <- getFitEsts(myfit, newdata = NULL, p=c(0.025, 0.05, 0.25, 0.5, 0.75, 0.95, 0.975)) #
```

The median is `r interqs[4]` days and the  0.95 at `r interqs[6]`. 

Here is a plot of the estimated distribution together with the empirical survival curve from the data. 

```{r}
spdays <- seq(0,20, by=0.05)
spdensity <- dweibull(days, shape = exp(spfit$coefficients[1]), scale = exp(spfit$coefficients[2]))

ggsp = ggsurvplot(
  fit=survfit(Surv(spdata$minIncTimes, spdata$maxIncTimes, type="interval2")~1, data=spdata), 
  xlab="Days",  ylab = "Overall probability of no symptoms yet")
pdata <- data.frame(days=days,  fitsurv=1-pweibull(days, shape = exp(myfit$coefficients[1]),
                      scale = exp(myfit$coefficients[2])))
ggsp$plot + geom_line(data = pdata, aes(x = spdays, y = fitsurv))
```


## Serial interval 
We will estimate the serial interval using the 'interval case to case' approach given in Vink et al (https://academic.oup.com/aje/article/180/9/865/2739204). 

The dataset has several instances where a putative infector or contact is known. These are listed in the 'related_cases' column. We first make a graph in which nodes are individuals and edges are present from cases listed as possible sources, to the cases for whom they are possible sources. 

```{r}
spnodes <- spdata$CaseID

## How to extract caseIDs from related_cases column - there are multiple values in some cells, separated by commas
spdata$related_cases #7 max within one cell

# Split into separate columns
spdata <- separate(spdata,
                    col = related_cases,
                    into = paste("contactID", 1:7, sep = "_"),
                    fill = "right")

# Turn into numeric values
spdata <- mutate(spdata, 
                    contactID_1 = as.numeric(contactID_1),
                    contactID_2 = as.numeric(contactID_2),
                    contactID_3 = as.numeric(contactID_3),
                    contactID_4 = as.numeric(contactID_4),
                    contactID_5 = as.numeric(contactID_5),
                    contactID_6 = as.numeric(contactID_6),
                    contactID_7 = as.numeric(contactID_7))

# Select down to columns of interest
spedges <- select(spdata, c(CaseID, starts_with("contactID")))

# Remove rows with NAs for at least one contact
spedges <- filter(spedges, !is.na(spedges$contactID_1)) #43 CasesIDs with 1 or more possible contacts
```

From this edge list we can use visNetwork to visualise the graph. Make 'group' based on source of probably infection. Colours are from the infection source column (but we should have a better colour scheme, like date of symptom onset). 

```{r}
# Turn 'presumed_reason' into lower case and get trim any whitespace so don't have issues with case sensitivity, etc
spdata$presumed_reason <- str_to_lower(spdata$presumed_reason)
spdata$presumed_reason <- str_trim(spdata$presumed_reason)
table(spdata$presumed_reason)
sum(is.na(spdata$presumed_reason)) #15 NAs

# Make a new column where we group the 'presumed_reason' under a label (known relationship, gathering, wuhan travel) for each of the above three groups
spdata <- mutate(spdata, presumed_reason_group = case_when(!is.na(str_match(presumed_reason, "symptom onset|via")) ~ "known relationship",
                                                           !is.na(str_match(presumed_reason, "grace")) ~ "grace assembly of god",
                                                           !is.na(str_match(presumed_reason, "grand")) ~ "grand hyatt singapore",
                                                           !is.na(str_match(presumed_reason, "life")) ~ "life church",
                                                           !is.na(str_match(presumed_reason, "seletar")) ~ "seletar aerospace heights",
                                                           !is.na(str_match(presumed_reason, "yong")) ~ "yong thai hang",
                                                           !is.na(str_match(presumed_reason, "wuhan|airport")) ~ "wuhan travel", #'airport' case (CaseID 17) does not have 'wuhan' in reason but does have it under 'Case' column that they are from Wuhan
                                                           is.na(presumed_reason) ~ "unknown",
                                                           TRUE ~ "other")) #should not be any other, so is just a double check this has run correctly, especially as dataset grows
table(spdata$presumed_reason_group)

spedges$arrows <- "to"
nodes.df <- data.frame(id=spdata$CaseID, label=spdata$CaseID,
                   group=spdata$presumed_reason_group)
#Ensure dfs are the same size
nodes.df <- filter(nodes.df, id %in% spedges$CaseID)
glimpse(nodes.df)
visNetwork(nodes.df, spedges)

#Hmmmm...this is not looking right...
```
