---
title: "Cluster Network Data Preprocessing & Plotting"
author: "Emma Garlock; Yen-Hsiang (Brian) Lee"
date: "18/02/2020"
updated: "23/02/2020"
output: html_document
---

**Set the working directory to be EpiCoronaHack_Cluster/Clustering/network_diagram**

Load all the libraries we will need 
```{r}
# getwd()
library(ggplot2)
library(tidyverse)
library(ggnetwork)
library(OneR)
library(igraph)
library(here)
library(viridis)
library(paletteer)
```

load up all the links we have (just the first two column of the dataframe expanded)
```{r}
ncov = read.csv("../data/COVID-19_Singapore.csv")
# ncov=read.csv("../../Data/COVID-19_Singapore.csv")
ncov_sing = ncov[ncov$country == "Singapore", ]
names(ncov_sing)[1] <- 'CaseID'
```


```{r}
links_two = ncov_sing[, c(1:3)]
links_condense = gather(links_two, key="linktype", value="relations", Related.cases:Cluster.links)
```

```{r}
links_exp = separate(links_condense, 'relations', paste("relations", 1:21, sep=""), sep=",", extra="drop")
edges = gather(links_exp, rel, link, relations1:relations21) %>% select("CaseID", "link", "linktype")
# write.csv(edges,("../data/links_long.csv"))
```
make it clear that those links are the edges we will be using. Also omit any of the cases that don't have  links to others in the dataset 
```{r}
names(edges) = c("from", "to", "linktype")
edges$from = as.character(edges$from)
edges$from = as.integer(edges$from)
edges$to = as.character(edges$to)
edges$to = as.integer(edges$to)
edges_reduced = edges[!(is.na(edges$to) | edges$to == "" | edges$to == " "), ]

```
use this chunk to get rid of duplicate links...somehow
```{r}
edges_reduced[1:2] <- t(apply(edges_reduced[1:2], 1, sort))
edges_dist = select(edges_reduced, from, to)
edges_dist = distinct(edges_dist)

```

Load the node info. This basically just supplies the metadata about each of the nodes in the edges df. 
```{r}
nodes_sing = ncov_sing
```
This chunk does not always have to be run, only if you need to generate a new set of coordinates for the network diagram. If you skip to the chunk below there is a step where you can load n, a df that already has the coordinates you will need 
```{r}
cov_net = graph_from_data_frame(edges_dist, vertices = nodes_sing, directed = FALSE)
n = ggnetwork(cov_net)
# write.csv(n, "../data/cov_net.csv")
# saveRDS(n, "../data/cov_net.rds")
```

Load the dataset that has the coordinates mapped out, also convert the age coumn to an intger so that we can bin it for nice plotting.
```{r}
# n = read.csv("../data/cov_net.csv")
n$age = as.character(n$age)
n$age = as.integer(n$age)
n$age_bin = bin(n$age, nbins = 5)
write.csv(n, "../data/COVID-19_Singapore_Network_plot.csv")
```
Make the plot! 
We can subset by 

 * age (binned)
 * cluster 
 * hospital 
 * outcomes 
 * travel_history_location 
```{r}

cov_net = ggplot(n, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_edges(aes(), subset(n, Related.cases != ""), linetype="solid") +
  geom_nodes(aes(color=cluster),size=8) +
  geom_text(aes(label=name), check_overlap = TRUE) +
  ggtitle("COVID-19 Singapore Cluster Network") +
  guides(colour = guide_legend(title.position="top", title.hjust = 0.5, override.aes = list(size=3))) +
  scale_colour_viridis_d(na.value="lightgrey") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) + #centre main title
  theme(panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.text=element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal")

 
cov_net

```
#####THIS CODE IS A MESS RIGHT NOW-working to fix this right now but I wanted to push it########

Tianjin Data
```{r}
tj=read.csv(here("Clustering/Tianjin135casesFeb22.csv"))
edges_unclean=tj[,c(1,7)]
edges_unclean$close_conn=str_extract_all(edges_unclean$Infection_source, regex("TJ\\d+"))
edges_unclean$close_conn=ifelse(edges_unclean$close_conn=="character(0)",as.character(edges_unclean$Infection_source),edges_unclean$close_conn)
edges_unclean$close_conn=gsub("^c\\(|\\)$", "", edges_unclean$close_conn)


write.csv(edges_unclean,here("Clustering/tianjin_edges.csv"))
```
Using the edges from the Tianjin.Rmd cluster 
```{r}
edges[1:2] <- t(apply(edges[1:2], 1, sort))
edges_dist = select(edges, from, to)
edges_dist = distinct(edges_dist)

edges_full=read.csv(here("Clustering/tianjin_edges_clean.csv"))
edges_cleaned=edges_full[,c(2,4)]
names(edges_cleaned)=c("name","cluster")



```

```{r}

tj_net=graph_from_data_frame(edges_dist,vertices = tdata,directed = FALSE)
ntj=ggnetwork(tj_net)

ntj_clean=merge(ntj,edges_cleaned,by="name")
ntj_clean$case_num=str_remove_all(ntj_clean$name,"[TJ]")


```
```{r}
tj_cov_net = ggplot(ntj_clean, aes(x = x, y = y, xend = xend, yend = yend))+
  geom_edges(aes(), linetype="solid",colour="black") +
  geom_nodes(aes(color=cluster),size=8) +
  geom_text(aes(label=case_num), check_overlap = TRUE)+
  scale_color_viridis_d()+
  ggtitle("Clusters of COVID-19 cases in Tianjin, China")+
  theme_blank()

tj_cov_net

```

```{r}
edges[1:2] <- t(apply(edges[1:2], 1, sort))
edges_dist = select(edges, from, to)
edges_dist = distinct(edges_dist)
```

```{r}
edges_cleaned$cluster=str_replace(edges_cleaned$cluster,"coworker ","coworker")
edges_cleaned$cluster=str_replace(edges_cleaned$cluster,"confirmed patient","unknown")
edges_cleaned$cluster=str_replace(edges_cleaned$cluster,"not specified","unknown")

cluster_list=(unique(edges_cleaned$cluster))



```
```{r}
for (c in cluster_list){
  assign(paste0("list",c),edges_cleaned[edges_cleaned$cluster==c,1])
}


```
```{r}
edges_full$cluster_list=ifelse(edges_full$infection_source_named=="Wuhan",list(as.character(listWuhan)),
                            ifelse(edges_full$infection_source_named=="Wuhan; train import",list(as.character(`listWuhan; train import`)),
                            ifelse(edges_full$infection_source_named=="unkown",list(as.character(listunknown)),
                            ifelse(edges_full$infection_source_named=="travel",list(as.character(`listtravel `)),
                            ifelse(edges_full$infection_source_named=="train",list(as.character(listtrain)),
                            ifelse(edges_full$infection_source_named=="mall",list(as.character(listmall)),
                            ifelse(edges_full$infection_source_named=="Hubei",list(as.character(listHubei)), 
                            ifelse(edges_full$infection_source_named=="coworker",list(as.character(listcoworker)),
                            ifelse(edges_full$infection_source_named==" close contact ",list(as.character(`list close contact `)),NA
                                   )))))))))
edges_full$contact_list=paste(edges_full$edge1,edges_full$edge2,edges_full$edge3,edges_full$edge4,sep=",")

edges_reduced=edges_full[,c(2,10,11)]
edges_reduced$edge_list=as.character(edges_reduced$edge_list)
edges_reduced$edge_list=gsub("^c\\(|\\)$", "", edges_reduced$edge_list)
edges_reduced$edge_list=gsub('"', "", edges_reduced$edge_list)

edges_long=gather(edges_reduced,type,links,edge_list,contact_list)
edges_long=separate(edges_long$links, paste("relations", 1:41, sep=""),sep=",",extra="drop")
```

                         
