---
title: "Network data cleaning and plotting"
author: "Emma Garlock"
date: "2/18/2020"
output: html_document
---

#Set the working directory to be EpiCoronaHack_Cluster/Clustering/network_diagram
Load all the libraries we will need 
```{r}
library(ggplot2)
library(tidyverse)
library(ggnetwork)
library(OneR)
library(igraph)
library(here)
```

load up all the links we have (just the first two column of the dataframe expanded)
```{r}
ncov=read.csv(here("./data/singapore_ncov_2019.csv"))
# ncov=read.csv("../../Data/nCov_Singapore_2019.csv")
ncov_sing=ncov[ncov$country=="Singapore",]
names(ncov_sing)[1] <- 'CaseID'
```

```{r}
links_condense=ncov_sing[,c(1,2)]
links_exp=separate(links_condense, 'Related.cases', paste("relations", 1:7, sep=""), sep=",", extra="drop")
links_long=gather(links_exp,rel,link,relations1:relations7)%>%select("CaseID","link")
write.csv(links_long,("../data/links_long.csv"))
```
make it clear that those links are the edges we will be using. Also omit any of the cases that don't have  links to others in the dataset 
```{r}
edges=links_long
names(edges)=c("from","to")
edges$from=as.character(edges$from)
edges$from=as.integer(edges$from)
edges$to=as.character(edges$to)
edges$to=as.integer(edges$to)
edges_reduced=edges[!(is.na(edges$to) | edges$to==""| edges$to==" "), ]
```

Load the node info. This basically just supplies the metadata about each of the nodes in the edges df. 
```{r}
nodes_sing=ncov_sing
```
This chunk does not always have to be run, only if you need to generate a new set of coordinates for the network diagram. If you skip to the chunk below there is a step where you can load n, a df that already has the coordinates you will need 
```{r}
cov_net=graph_from_data_frame(edges_reduced,vertices = nodes_sing,directed=TRUE)
n=ggnetwork(cov_net)
write.csv(n,"../data/good_coords.csv")
saveRDS(n, "../data/cov_net.rds")
```

Load the dataset that has the coordinates mapped out, also convert the age coumn to an intger so that we can bin it for nice plotting.
```{r}
n=read.csv("../data/good_coords.csv")
n$age=as.character(n$age)
n$age=as.integer(n$age)
n$age_bin=bin(n$age,nbins=5)
```
Make the plot! 
We can subset by 
* age (binned)
* cluster 
* hospital 
* outcomes 
* travel_history_location 
```{r}

cov_net=ggplot(n)+
  geom_edges(aes(x = x, y = y, xend = xend, yend = yend),arrow = NULL)+
  geom_nodes(aes(x = x, y = y,color=cluster),size=8)+
  geom_text(aes(x = x, y = y,label=name),check_overlap = TRUE)+
  guides(colour = guide_legend(title.position="top",title.hjust = 0.5,override.aes = list(size=3)))+
  theme_bw()+
  theme(panel.grid = element_blank(),
        axis.title = element_blank(),
        axis.line = element_blank(),
        axis.text=element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_blank(),
        legend.position = "bottom",
        legend.direction = "horizontal")
 
cov_net

```



